---
title: "bunbun's Finals Econometric"
output: 
  flexdashboard::flex_dashboard:
    theme : flatly
    orientation: columns
    vertical_layout: fill
    source_code : NULL
    
---

```{r setup, include=FALSE}
library(flexdashboard)
library(jsonlite)
library(dplyr)
library(tidyr)
library(highcharter)
library(gp)
library(data.table)
library(DT)

```

# Data 

## Column {.tabset .tabset-pills}


### Data

The data is about Indonesia's Economic year 2010-2021.

```{r}
library(readxl)
setwd(getwd())

interest_rate <- read_excel("data_UAS/INTEREST_RATE.xlsx")
stock_index <- read.csv("data_uas/JKLQ45_uas.csv")
consumer_confidence <- read_excel("data_UAS/Book1.xlsx", sheet= "kepercayaan_konsumen")
inflation_rate <- read_excel("data_UAS/Book1.xlsx", sheet= "tingkat_inflasi")
Kurs_USD <- read_excel("data_UAS/Kurs_Transaksi_USD.xlsx")

library(reshape2)
library(dplyr)
interest_rate <- melt(interest_rate, "Bulan/Tahun") 
interest_rate <- interest_rate%>%
  arrange(`Bulan/Tahun`)


consumer_confidence <- melt(consumer_confidence, "kepercayaan konsumen")
inflation_rate <- melt(inflation_rate, "tingkat inflasi")

dc <- data.frame(
  dates = seq(as.Date("2010/1/1"),as.Date("2021/12/1"), "month"),
  interest_rate = interest_rate$value,
  stock_index = stock_index$Close,
  consumer_confidence = consumer_confidence$value,
  inflation_rate = inflation_rate$value
)

unemployment_rate <- read_excel("data_UAS/Pengangguran.xlsx", sheet="Sheet2")
ur <- unemployment_rate$unemployment_rate

ali <- data.frame(
   0.1909188,
   0.3676955,
   0.1697056,
   0.205061,
   0.1697056,
   0.2616295,
   0.07778175,
   0.1202082,
   0.1414214,
   0.1767767,
   1.506137
)

set.seed(1993)

a= c(
rnorm(6, mean= ur[1] , sd = ali[1,1]),
rnorm(6, mean= ur[2] , sd = ali[1,1]),
rnorm(6, mean= ur[3] , sd = ali[1,2]),
rnorm(6, mean= ur[4] , sd = ali[1,2]),
rnorm(6, mean= ur[5] , sd = ali[1,3]),
rnorm(6, mean= ur[6] , sd = ali[1,3]),
rnorm(6, mean= ur[7] , sd = ali[1,4]),
rnorm(6, mean= ur[8] , sd = ali[1,4]),
rnorm(6, mean= ur[9] , sd = ali[1,5]),
rnorm(6, mean= ur[10], sd = ali[1,5]),
rnorm(6, mean= ur[11], sd = ali[1,6]),
rnorm(6, mean= ur[12], sd = ali[1,6]),
rnorm(6, mean= ur[13], sd = ali[1,7]),
rnorm(6, mean= ur[14], sd = ali[1,7]),
rnorm(6, mean= ur[15], sd = ali[1,8]),
rnorm(6, mean= ur[16], sd = ali[1,8]),
rnorm(6, mean= ur[17], sd = ali[1,9]),
rnorm(6, mean= ur[18], sd = ali[1,9]),
rnorm(6, mean= ur[19], sd = ali[1,10]),
rnorm(6, mean= ur[20], sd = ali[1,10]),
rnorm(6, mean= ur[21], sd = ali[1,11]),
rnorm(6, mean= ur[22], sd = ali[1,11])
)

c <- c(rnorm(12, mean= mean(ur), sd = sd(ur)) )

dc$unemployment_rate = c(a,c)

percentage <- c(5.69,6.17,5.8,6.9,
                  6.48,6.27,6.01,5.94,
                  6.11,6.21,5.94,5.87,
                  5.54,5.59,5.52,5.58,
                  5.12,4.94,4.93,5.05,
                  4.83,4.74,4.78,5.15,
                  4.94,5.21,5.03,4.94,
                  5.01,5.01,5.06,5.19,
                  5.06,5.27,5.17,5.18,
                  5.07,5.05,5.02,4.97,
                  2.97,-5.32,-3.49,-2.19,
                  -0.69,7.08,3.53,5.03)
sd_gdp <- c(
 sd(c(5.69,6.17,5.8,6.9)),
 sd(c(6.48,6.27,6.01,5.94)),
 sd(c(6.11,6.21,5.94,5.87)),
 sd(c(5.54,5.59,5.52,5.58)),
 sd(c(5.12,4.94,4.93,5.05)),
 sd(c(4.83,4.74,4.78,5.15)),
 sd(c(4.94,5.21,5.03,4.94)),
 sd(c(5.01,5.01,5.06,5.19)),
 sd(c(5.06,5.27,5.17,5.18)),
 sd(c(5.07,5.05,5.02,4.97)),
 sd(c(2.97,-5.32,-3.49,-2.19)),
 sd(c(-0.69,7.08,3.53,5.03)))
 
GDP_growth <- c(
rnorm(3, mean = percentage[1],  sd= sd_gdp[1]),
rnorm(3, mean = percentage[2],  sd= sd_gdp[1]),
rnorm(3, mean = percentage[3],  sd= sd_gdp[1]),
rnorm(3, mean = percentage[4],  sd= sd_gdp[1]),
rnorm(3, mean = percentage[5],  sd= sd_gdp[2]),
rnorm(3, mean = percentage[6],  sd= sd_gdp[2]),
rnorm(3, mean = percentage[7],  sd= sd_gdp[2]),
rnorm(3, mean = percentage[8],  sd= sd_gdp[2]),
rnorm(3, mean = percentage[9],  sd= sd_gdp[3]),
rnorm(3, mean = percentage[10], sd= sd_gdp[3]),
rnorm(3, mean = percentage[11], sd= sd_gdp[3]),
rnorm(3, mean = percentage[12], sd= sd_gdp[3]),
rnorm(3, mean = percentage[13], sd= sd_gdp[4]),
rnorm(3, mean = percentage[14], sd= sd_gdp[4]),
rnorm(3, mean = percentage[15], sd= sd_gdp[4]),
rnorm(3, mean = percentage[16], sd= sd_gdp[4]),
rnorm(3, mean = percentage[17], sd= sd_gdp[5]),
rnorm(3, mean = percentage[18], sd= sd_gdp[5]),
rnorm(3, mean = percentage[19], sd= sd_gdp[5]),
rnorm(3, mean = percentage[20], sd= sd_gdp[5]),
rnorm(3, mean = percentage[21], sd= sd_gdp[6]),
rnorm(3, mean = percentage[22], sd= sd_gdp[6]),
rnorm(3, mean = percentage[23], sd= sd_gdp[6]),
rnorm(3, mean = percentage[24], sd= sd_gdp[6]),
rnorm(3, mean = percentage[25], sd= sd_gdp[7]),
rnorm(3, mean = percentage[26], sd= sd_gdp[7]),
rnorm(3, mean = percentage[27], sd= sd_gdp[7]),
rnorm(3, mean = percentage[28], sd= sd_gdp[7]),
rnorm(3, mean = percentage[29], sd= sd_gdp[8]),
rnorm(3, mean = percentage[30], sd= sd_gdp[8]),
rnorm(3, mean = percentage[31], sd= sd_gdp[8]),
rnorm(3, mean = percentage[32], sd= sd_gdp[8]),
rnorm(3, mean = percentage[33], sd= sd_gdp[9]),
rnorm(3, mean = percentage[34], sd= sd_gdp[9]),
rnorm(3, mean = percentage[35], sd= sd_gdp[9]),
rnorm(3, mean = percentage[36], sd= sd_gdp[9]),
rnorm(3, mean = percentage[37], sd= sd_gdp[10]),
rnorm(3, mean = percentage[38], sd= sd_gdp[10]),
rnorm(3, mean = percentage[39], sd= sd_gdp[10]),
rnorm(3, mean = percentage[40], sd= sd_gdp[10]),
rnorm(3, mean = percentage[41], sd= sd_gdp[11]),
rnorm(3, mean = percentage[42], sd= sd_gdp[11]),
rnorm(3, mean = percentage[43], sd= sd_gdp[11]),
rnorm(3, mean = percentage[44], sd= sd_gdp[11]),
rnorm(3, mean = percentage[45], sd= sd_gdp[12]),
rnorm(3, mean = percentage[46], sd= sd_gdp[12]),
rnorm(3, mean = percentage[47], sd= sd_gdp[12]),
rnorm(3, mean = percentage[48], sd= sd_gdp[12])
)
dc$GDP_growth = GDP_growth
GDP_growth<- data.frame(GDP_growth)

library(dplyr)
library(tidyr)
USD_buy <- Kurs_USD%>%
  separate(Tanggal, c("bulan","hari", "tahun"))%>%
  group_by(tahun,bulan)%>%
  summarise(BuyUSD = mean(`Kurs Jual`))%>%
  mutate_at(c("tahun","bulan"), as.numeric)%>%
  arrange(tahun,bulan)
dc$USD = USD_buy$BuyUSD
DT::datatable(dc)
```


### Error Time Series Prediction

```{r}
library(forecast)
dc_arima1 <- arima(dc$USD, order = c(1,1,1)) # ARIMA(1,1,1)
dc_arima2 <- arima(dc$USD, order = c(2,1,1)) # ARIMA(2,1,1)
dc_arima3 <- arima(dc$USD, order = c(3,1,1)) # ARIMA(3,1,1)
dc_arima4 <- arima(dc$USD, order = c(4,1,1)) # ARIMA(4,1,1)

auto_arima <- auto.arima(dc$USD, max.p = 4, d = 1, max.q = 1) 


ts_error <-data.frame(
  actual = dc$USD
)
ts_error$"ARIMA(1,1,1)" = dc_arima1$residuals 
ts_error$"ARIMA(2,1,1)" = dc_arima2$residuals 
ts_error$"ARIMA(3,1,1)" = dc_arima3$residuals 
ts_error$"ARIMA(4,1,1)" = dc_arima4$residuals 
ts_error$"ARIMA(2,1,0)" = auto_arima$residuals 

library(rugarch)
# ARCH(1) model
arch1_spec <- ugarchspec(variance.model= list(model="sGARCH", garchOrder = c(1,0)), mean.model = list(armaOrder = c(0,0), include.mean=T))
arch1_fit <- ugarchfit(spec=arch1_spec, data=dc$USD)

ts_error$"ARCH(1)" = residuals(arch1_fit)

# GARCH(1) model
garch1_spec <- ugarchspec(variance.model= list(model="sGARCH", garchOrder = c(1,1)), mean.model = list(armaOrder = c(0,0), include.mean=T))
garch1_fit <- ugarchfit(spec=garch1_spec, data=dc$USD)

ts_error$"GARCH(1,1)" = residuals(garch1_fit)

PE_ts <-abs(ts_error)/ ts_error$actual
PE_ts_tes <-PE_ts[c(116:144),] 

acc_ts <- data.frame(
  Method = colnames(PE_ts_tes[,-1]),
  MAPE = c(mean(PE_ts_tes$`ARIMA(1,1,1)`) *100,
           mean(PE_ts_tes$`ARIMA(2,1,1)`) *100,
           mean(PE_ts_tes$`ARIMA(3,1,1)`) *100,
           mean(PE_ts_tes$`ARIMA(4,1,1)`) *100,
           mean(PE_ts_tes$`ARIMA(2,1,0)`) *100,
           mean(PE_ts_tes$`ARCH(1)`)*100,
           mean(PE_ts_tes$`GARCH(1,1)`*100)
  )
)

library(fpp2)
holt_train <- window(dc$USD, end = 115)
holt_test <- window(dc$USD, start = 116)

beta <- seq(.0001, .5, by = .001)
RMSE <- NA
for(i in seq_along(beta)) {
  fit <- holt(holt_train,
              beta = beta[i], 
              h = 100)
  RMSE[i] <- accuracy(fit, 
                      holt_test)[2,2]
}

beta.fit <- data_frame(beta, RMSE)
beta.min <- filter(beta.fit, 
                   RMSE == min(RMSE))

holts <- holt(holt_train, h=100)
holts2 <- holt(holt_train, h=100, beta=beta.min$beta)

holts_acc <- data.frame(
  Method = c("Holt's b=0.0568", "Holt's b=0.0451"),
  MAPE = c(1.957945, 2.068370)
  )

acc_ts <- rbind(acc_ts, holts_acc)
datatable(acc_ts)
```


# Area Dist

## Column { .tabset .tabset-pills data-width=500}




### Rate
Interest Rate, Inflation Rate, Unemployment Rate and GDP Growth
```{r}
hchart(density(dc$interest_rate), type = "area", color= "#003285",name = "interest rate")%>%
  hc_add_series(density(dc$inflation_rate), type="area", color = "#FF7F3E", name="inflation rate")%>%
  hc_add_series(density(dc$unemployment_rate), type="area", color = "#94FFD8", name="unemployment rate")%>%
  hc_add_series(density(dc$GDP_growth), type="area", color = "#FF0080", name="GDP growth")%>%
  hc_title(text = "Density Plots of Rate",
           style = list(color = "#240750", useHTML = TRUE))
```

### Stock Index

```{r}
hchart(density(dc$stock_index), type="area", color="#A0DEFF", name="LQ45 Stock Index")%>%
  hc_title(text = "Density Plots of Stock Index LQ45",
           style = list(color = "#240750", useHTML = TRUE))
```


### Exchange Rates
```{r}
hchart(density(dc$USD), type="area", color="#A1DD70", name="USD to Rupiah") %>%
  hc_title(text = "Density Plots of USD",
           style = list(color = "#240750", useHTML = TRUE))
```


## Column 

### Correlation Plot

```{r}
library(plotly)
fig <- plot_ly(x = colnames(dc[,-1]), y = colnames(dc[,-1]), z =cor(dc[,-1]), type = "heatmap")%>%
        layout(title = "Correlation between variables")

fig
```



# Time Series 


## Column {.tabset .tabset-pills}

### Prediction

```{r}
ts_dc <- ts((dc$USD), start = c(2010,1), end = c(2021,12), frequency = 12)
decomp <- decompose(ts_dc)
ts <- ts(c(dc$USD), start = c(2010,1), 
                frequency = 12) 
ts2 <- ts(c(fitted.values(dc_arima3),14380.68,14370.20), start = c(2010,1), 
                frequency = 12) 
hchart(ts2, type="line", name="Predicted US Dollar", color = "#FF9A00")%>%
  hc_add_series(ts, type="line", name= "Actual US Dollar", color="#615EFC")%>%
  hc_title(text="Actual vs Prediction Rp to USD by ARIMA(3,1,1)")%>%
  hc_add_series(decomp$trend, type="line", name="trend", color="#A1DD70")
```


### Box-Jenkins

```{r}
p1= predict(dc_arima1,2)
p2= predict(dc_arima2,2)
p3= predict(dc_arima3,2)
p4= predict(dc_arima4,2)

arima5 <- arima(dc$USD, order = c(2,1,0))
p5= predict(arima5,2)

ts <- ts(c(dc$USD), start = c(2010,1), frequency = 12) 

a1 = ts(c(fitted.values(dc_arima1) ,p1$pred), start = c(2010,1), frequency = 12) 
a2 = ts(c(fitted.values(dc_arima2) ,p2$pred), start = c(2010,1), frequency = 12) 
a3 = ts(c(fitted.values(dc_arima3) ,p3$pred), start = c(2010,1), frequency = 12) 
a4 = ts(c(fitted.values(dc_arima4) ,p4$pred), start = c(2010,1), frequency = 12) 
a5 = ts(c(fitted.values(arima5),p5$pred), start = c(2010,1), frequency = 12) 



hchart(ts, type="line", name="Actual US Dollar", color = "black")%>%
  hc_add_series(a1, type="line", name= "ARIMA(1,1,1)", color="#03AED2")%>%
  hc_add_series(a2, type="line", name= "ARIMA(2,1,1)", color="#68D2E8")%>%
  hc_add_series(a3, type="line", name= "ARIMA(3,1,1)", color="#FDDE55")%>%
  hc_add_series(a4, type="line", name= "ARIMA(4,1,1)", color="#615EFC")%>%
  hc_add_series(a5, type="line", name= "ARIMA(2,1,0)", color="#A3D8FF")%>%
  hc_title(text="Prediction Rp to USD by Box-Jenkins")

```


### Holts

```{r}
h1= predict(holts,2)
h2=  predict(holts2,2)

h1= data.frame(h1)
h2 = predict(h2)
h11 = ts(c(fitted.values(holts), h1$Point.Forecast[1:30]	), start = c(2010,1), frequency = 12) 
h22 = ts(c(fitted.values(holts2), h2$Point.Forecast[1:30]), start = c(2010,1), frequency = 12) 


hchart(ts, type="line", name="Actual US Dollar", color = "black")%>%
  hc_add_series(h11, type="line", name= "Holt's b=0.0568", color="#F9A828")%>%
  hc_add_series(h22, type="line", name= "Holt's b=0.0451", color="#F29191")%>%
  hc_title(text="Prediction Rp to USD by Holt's Method")

```


### Combined Plot

```{r}

hchart(ts, type="line", name="Actual US Dollar", color = "black")%>%
  hc_add_series(h11, type="line", name= "Holt's b=0.0568", color="#F9A828")%>%
  hc_add_series(a1, type="line", name= "ARIMA(1,1,1)", color="#03AED2")%>%
  hc_add_series(a2, type="line", name= "ARIMA(2,1,1)", color="#68D2E8")%>%
  hc_add_series(a3, type="line", name= "ARIMA(3,1,1)", color="#FDDE55")%>%
  hc_add_series(a4, type="line", name= "ARIMA(4,1,1)", color="#615EFC")%>%
  hc_add_series(a5, type="line", name= "ARIMA(2,1,0)", color="#A3D8FF")%>%  
  hc_add_series(h22, type="line", name= "Holt's b=0.0451", color="#F29191")%>%
  hc_title(text="Prediction Rp to USD Time Series Model")
```



```{r eval=FALSE}
### ARCH GARCHar1 = predict(arch_fit,2)
ar2 =  predict(arch2_fit,2)

ar1= data.frame(h1)
ar2 = data.frame(h2)
arch_fit
ar11 = ts(c(gamma(arch_fit), ar1$Point.Forecast[1:2]	), start = c(2010,1), frequency = 12) 
ar22 = ts(c(fitted.values(arch2_fit), ar2$Point.Forecast[1:2]), start = c(2010,1), frequency = 12) 


hchart(ts, type="line", name="Actual US Dollar", color = "black")%>%
  hc_add_series(ar11, type="line", name= "Holt's b=0.0568", color="#F9A828")%>%
  hc_add_series(ar22, type="line", name= "Holt's b=0.0451", color="#F29191")%>%
  hc_title(text="Prediction Rp to USD by Holt's Method")

```

# Reference

+ Makridakis, S., Wheelwright, S. C., & Hyndman, R. J. (2008). Forecasting methods and applications. John wiley & sons.

+ https://www.datacamp.com/tutorial/tutorial-ridge-lasso-elastic-net

+ https://sscc.wisc.edu/sscc/pubs/RegDiag-R/linearity.html

+ https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/seq.Date

+ https://www.geeksforgeeks.org/decision-tree-for-regression-in-r-programming/

+ https://www.geeksforgeeks.org/random-forest-approach-for-regression-in-r-programming/

+ https://www.geeksforgeeks.org/exponential-smoothing-in-r-programming/

+ https://colorhunt.co/

